\subsection{Poisson Distribution and Process}
\label{sec:Method:Poisson}
The Poisson distribution models the probability that a given number of independent events, from a discrete random variable, will happen in a specific time interval \cite{PoissonScience}. 
The events will happen at some rate, defined by the symbol $\lambda$, also named the intensity function. 
For instance, the Poisson distribution could model the discrete random variable that represents two nodes interacting, which can only take two values either 1 (interaction) or 0 (no interaction).
More formally the probability mass function of the Poisson distribution \cite{Bas2019AProcess} (p. 136) is described by:

\begin{equation}
    p_k = \frac{e^{-\lambda}\lambda^k}{k!} \;\; \text{for} \;\; \lambda > 0 \;\; \text{and} \;\; k = 0,1,\dots
\end{equation}
Here, $k$ is the number of occurrences for a given event, in this case a node pair interacting and $p_k$ denotes the probability of getting $k$ occurrences.
Worth noting is that a random variable $X$, that is Poisson distributed with intensity function $\lambda$, has $E[X] = \lambda = Var[X]$.
Further, the intensity function $\lambda$ might be constant throughout time or it may be time-dependent as a function of time, ie. $\lambda(t)$.


\subsubsection{Poisson Point Process}
\label{sec:Method:Poisson:PoissonPointProcess}
In order to understand the Poisson point process, it is useful to first define the point process. 
A point process is a stochastic process with a collection of random variables that represent the arrival times of events i.e. when events happen. 
In the case of dynamic networks modelled in this project, this would be the time points when nodes interact. 
These arrival times can be defined as ${t_n ; n \ge 0}$, such that:

$$
t_0 = 0 < t_1 < t_2 < ... < t_n
$$
where $t_n$ is the arrival time of the \textit{n}th event.
\\\\
Each of these arrival times will happen one after another, and the times between them are called interarrival times. 
The interarrival times are independent random variables \cite{Bas2019AProcess} (p. 132) and can be defined for the point process as:

\begin{align*}
T_1 &= t_1 - t_0 \\
T_2 &= t_2 - t_1 \\
&\dots \\
T_n &= t_n - t_{n-1}
\end{align*}
where ${T_n ; n \ge 1}$ is a stochastic process with the random variables denoting the interarrival times.
\clearpage
A Poisson point process with positive intensity function $\lambda > 0$ can be defined as a point process in which the interarrival times ${T_n ; n \ge 1}$ are independent exponentially distributed random variables. 
If the intensity function $\lambda$ of the Poisson process is constant for all time intervals, the process is said to be homogeneous. 
During the interarrival times, the Poisson process increments happen as:

\begin{align*}
X_1 &= X(t_1) - X(t_0) \\
X_2 &= X(t_2) - X(t_1) \\
&\;\;\;\;\;\;\;\dots \\
X_n &= X(t_n) - X(t_{n-1})
\end{align*}
These are independent random variables that yield the number of events happening in the time interval $(t_{n-1}, t_n]$. 
For the homogeneous Poisson process with $\lambda > 0$, with starting time $s \ge 0$ and an interarrival time step $t > 0$, the random variable $X(s+t) - X(s)$ is Poisson distributed, such that:

\begin{equation}
    P(X(s+t) - X(s) = k) = \frac{(\lambda)^k e^{-\lambda}}{k!} \;\; \text{for} \;\; k=0,1,\dots
\end{equation}
where $X(0) = 0$.
\\\\
If the intensity function of the Poisson process is not constant, ie. a function of time $\lambda(t)$, the process is said to be non-homogeneous or in-homogeneous \cite{Bas2019AProcess} (p. 136).
In the case of this project, the non-homogeneous Poisson point process is most relevant. 
This is because the intensity function, which describes the likelihood of node interactions, is governed by the squared Euclidean distance between the nodes. 
This distance changes through time, as the nodes change positions in Euclidean latent space, and as such the intensity function also changes over time.
\\
For the non-homogeneous Poisson process, the process increments with $X(s+t) - X(s)$ have ans has rate $\lambda(t)$, becoming Poisson distributed with parameter $\int_s^t \lambda(u)\; du$ \cite{Bas2019AProcess} (p. 136), such that:

\begin{equation}
    P(X(s+t) - X(s) = k) = \frac{\left(\int_s^{s+t} \lambda(u) \; du \right)^k exp\left(-\int_s^{s+t} \lambda(u) \; du\right)}{k!}
\end{equation}
This yield the probability of having $k$ events in the time interval $(s, s+t]$ for the non-homogeneous Poisson point process.
\clearpage
\subsubsection{Event Probability}
\label{sec:Method:Poisson:EventProbability}
Considering the case where the interarrival times $(s, s+t]$ are made infinitesimally small, such that $t = dt$, it will be assumed that no more than one event can occur during the interarrival time $(s, s+dt]$ \cite{Tommerup2021LearningNetworks}.
This leads to the following:

\begin{align}
P(X(s+dt) - X(s) = k) = 
\begin{cases}
    1 - \lambda(u) \; dt \; &\text{for} \; k=0 \\
    \lambda(u) \; dt \; &\text{for} \; k=1 \\
    0 \; &\text{for} \; k > 1
\end{cases}
\end{align}
This means that the event probability of a single event happening becomes:

\begin{align}
\begin{split}
    P(X(s+t) - X(s) = 1) 
    &= 
    \frac{\left(\int_s^{s+t} \lambda(u) \; du \right)^1 \exp \left(-\int_s^{s+t} \lambda(u) \; du\right)}{1!} \\
    &= 
    \left(\int_s^{s+t} \lambda(u) \; du \right) \exp \left(-\int_s^{s+t} \lambda(u) \; du\right)
\label{prob_single_event}
\end{split}
\end{align}
The above formula can be rewritten through the notion that for a function $f(x)$ with antiderivative $F(x)$ integrated over an infinitesimal interval it holds that:

\begin{equation}
    \int_s^{s+dt} f(u) \; du = F(u) \rvert_s^{s+dt} = F(s+dt) - F(s) = F^{\prime}(s) \; dst = f(s) \; dt
\end{equation}
Applied to (\ref{prob_single_event}), it yields:

\begin{equation}
    P(X(s+t) - X(s) = 1) = \lambda(s)dt \; e^{-\lambda(s)dt}
\end{equation}
This can be expanded with the power series, leading to the probability of an event happening at time $s$ being given by:

\begin{equation}
    P(X(s+t) - X(s) = 1) = \lambda(s) dt
\end{equation}
Here, the higher orders of the infinitesimal $ds$ are neglected.

\subsubsection{Non-Event Probability}
\label{sec:Method:Poisson:NonEventProbability}
A similar starting point can be taken for the non-event probability as with the event probability.
Considering the non-homogeneous Poisson point process with intensity $\lambda(t)$, the probability of no events happening during the interarrival time $(s, s+t]$ becomes:

\begin{equation}
    P(X(s+t) - X(s) = 0) 
    = 
    \frac{\left(\int_s^t \lambda(u)du \right)^0 \exp \left(- \int_s^t \lambda(u)du\right)}{0!}
    = 
    \exp \left(- \int_s^t \lambda(u)du\right)
\end{equation}
where $t = dt$ for infinitesimally small interarrival times \cite{Tommerup2021LearningNetworks}.
